---
title: "ML Final assignment"
output: html_document
---

Import data set and convert factor variables to numeric

```{r,message=FALSE,warning=FALSE,cache=TRUE}
library(dplyr,quietly = TRUE,warn.conflicts = FALSE)
library(caret,quietly = TRUE,warn.conflicts = FALSE)
library(gbm,quietly = TRUE,warn.conflicts = FALSE)
library(rpart,quietly = TRUE,warn.conflicts = FALSE)
library(rpart.plot,quietly = TRUE,warn.conflicts = FALSE)

pml.training <- read.csv("pml-training.csv")
pml.testing <- read.csv("pml-testing.csv", header=TRUE)
for (i in names(pml.training)[8:(length(pml.training)-1)]) {
    pml.training[[i]] <- as.numeric(as.character(pml.training[,i]))
}

```

Choose NA columns and store not NA columns in variable isnacolumn

Make sample data sets named "training" and "testing" without them

```{r,message=FALSE,warning=FALSE,cache=TRUE}
set.seed(1000)
isnacolumn<-sapply(pml.training, function(x){sum(is.na(x))<length(x)*0.1}) & 
    c(rep(FALSE,5),rep(TRUE,length(pml.training)-5))
inTrain <- createDataPartition(pml.training$classe, p = 3/4)[[1]]
training<-pml.training[inTrain,isnacolumn]
testing<-pml.training[-inTrain,isnacolumn]
```

First we use CART method

#CART method

```{r,message=FALSE,warning=FALSE,cache=TRUE}
fit<-train(classe~.,data=training,method="rpart")
confusionMatrix(testing$classe,predict(fit,testing))
prp(fit$finalModel)

```

Accuracy 0.57 is too small to stop finding

let's try to use GBM. We shall use PCA to reduce number of variables

#Gradient boosting mashine with PCA veriables reducing

```{r,message=FALSE,warning=FALSE,cache=TRUE}
set.seed(1000)
ctrl<-trainControl(verboseIter = FALSE,allowParallel = TRUE)
fit<-train(classe~.,data=training,method="gbm",preProcess = "pca", trControl=ctrl,verbose=FALSE)
confusionMatrix(testing$classe,predict(fit,testing))
summary(fit)
```

Accuracy of 0.83 is good, but let's try do not use PCA

#Gradient boosting mashine without PCA veriables reducing

```{r,message=FALSE,warning=FALSE,cache=TRUE}
set.seed(1000)
ctrl<-trainControl(verboseIter = FALSE,allowParallel = TRUE)
fit<-train(classe~.,data=training,method="gbm", trControl=ctrl,verbose=FALSE)
confusionMatrix(testing$classe,predict(fit,testing))
summary(fit)
```

Accuracy of 0.99 is best we can have.We do not need to use PCA 

Let's predict variables from test set

#Variables prediction

```{r,message=FALSE,warning=FALSE,cache=TRUE}
data.frame(x=pml.testing$X,prediction=predict(fit,pml.testing[,isnacolumn]))
```

